
## SPA coordinates {#spacoordinates}

### Summary {-}
The Cumulative Effects Framework (CEF), uses a list of seabird colonies or SPAs including coordinates representing their locations. Within the CEF, the SPA coordinates are given by a single point and all distances and flights are measured from this point. Where possible, the point location is taken from the SPA Site Details dataset (derived from Natura 2000 [@na_website_nd]) but some have been amended for this dataset. The distance by sea and flight path calculations require the point to be located in a grid cell designated as 'sea', so where the default point is in a 'land' cell, the SPA point has been moved to the nearest sea location. See CEF Documentation for further details.  

### Provenance {-}
The dataset was created initially from dataset 'SPA Site Details' (see Section 2.8). After comparison with the CEF spatial grid (see Section 6.3), points that were located in a 'land' cell were relocated to a 'sea' cell. Details of this are process are given in the CEF documentation.  

### Terms of Use {-}
Tbc

### Data set attribution {-}
You must cite: 

### Versions Available {-}
```{r spacoordinates, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}
library(readr)
library(knitr)
vers <- readr::read_csv('/data/DataStore/02_Colony/207_SPAcoords/vers_spacoordinates.csv', show_col_types = FALSE)
knitr::kable(vers)

```

#### Update information {-}
* july22 - Draft version

#### Data Store Location {-}
* Section: 02_Colony
* Folder: 207_SPAcoords
* Default version: july22

### Field Descriptions for version july22 {-}
```{r spacoordinatesfields, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}
fields <- readr::read_csv('/data/DataStore/02_Colony/207_SPAcoords/july22/fields_spacoordinates.csv', show_col_types = FALSE)
knitr::kable(fields)

```

### Supplemental information {-}

### Keywords {-}
SPA , coordinates, location

----

